---
title: "HW Day 1"
author: "Mai Ng"
date: "August 3, 2018"
output: html_document
---
## Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=TRUE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(lubridate)
```

```{r load-data, include=TRUE}
df <- read.csv("/Users/mainguyen/Downloads/MonthlyRainfall_mm_ByProvince.csv", header=TRUE, sep=",")
```

## Explore data 
```{r summary, inlude=TRUE}
summary(df)
```

```{r str, include=FALSE}
str(df)
```

```{r check-year, include=FALSE}
table(df$Year)
```

```{r check-province, include=FALSE}
table(df$Province)
```

```{r to-long, include=TRUE}
df_long <- gather(df, "month", "rain_fall", 3:9)

```

```{r check_month, }
table(df_long$month)
```

```{r reformat, include=TRUE}

## Reformate month
df_long$month <- factor(df_long$month, 
                        levels=c("May", "June", "July", "Aug.", "Sep.", "Oct.", "Nov."),
                        labels=c("May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov")
)

## Create a mon year var
df_long$mon_num <- ifelse(df_long$month == "May", "05-01", 
                          ifelse(df_long$month == "Jun", "06-01",
                                 ifelse(df_long$month == "Jul", "07-01", 
                                        ifelse(df_long$month == "Aug", "08-01",
                                               ifelse(df_long$month == "Sep", "09-01",
                                                      ifelse(df_long$month == "Oct", "10-01", 
                                                             ifelse(df_long$month == "Nov", "11-01", "Others"
                                                             )
                                                      )
                                               )
                                        )
                                 )
                          )
)
df_long$mon_year <- paste0(df_long$Year, "-", df_long$mon_num)
df_long$monyear <- format(as.Date(df_long$mon_year), "%Y-%m")
                              
## Refactor Provinces' names
df_long$Province <- factor(df_long$Province, levels=c("Lai Chau", "Tuyen Quang", "Son La", 
                                                      "Ha Noi", "Bai Chay", "Nam Dinh",
                                                      "Vinh", "Hue", "Da Nang", 
                                                      "Qui Nhon", "Playku", "Da Lat",
                                                      "Nha Trang", "Vung Tau", "Ca Mau")
)
```

## Visualizing

### Plot 1
**Seem like dry season is becoming drier as October now has less rain than before**

**In dry season, we also see a larger gap in precipitation among provinces**
```{r month-change, include=TRUE}
# color_palette <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
#  "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
#  "#920000","#924900","#db6d00","#24ff24","#ffff6d")
month_change <- ggplot(df_long, aes(x=Year, y=rain_fall, colour=Province)) + 
  geom_line() +
  #geom_point(size=1.5, shape=20) +
  #geom_hline(yintercept=100, color = "red", size=1.5) +
  facet_wrap(~month, ncol=7) +
  ggtitle("How does rainfall change over the years? - A view on climate change") +
  #theme(text = element_text(size = 16, family="Helvetica", face="italic")) + 
  theme_grey() +
  labs(y = "Precipitation", colour="Province") +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  theme(axis.text.x = element_text(size = 6)) +
  theme(axis.text.y = element_text(size = 6))
  #scale_colour_brewer(palette=color_palette)
month_change
```


### Plot 2
**As, I can see there should be some changes in rainfall, I take a look at the trend in total rainfall in each month over the years** 

**It is quite interesting to see following a month with much rain is a month with less rain and vice versa** 
```{r total-avg, inlude=TRUE}
df_long_cal <- df_long %>%
  group_by(Year, month) %>%
  summarise(total=sum(rain_fall),
          avg=mean(rain_fall))

total_change <- ggplot(df_long_cal, aes(x=Year, y=total)) + 
  geom_line() +
  facet_wrap(~month, ncol=7) +
  ggtitle("How does total rainfall change over the years? - A view on climate change") +
  #theme(text = element_text(size = 16, family="Helvetica", face="italic")) + 
  theme_grey() +
  labs(y = "Precipitation") +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  theme(axis.text.x = element_text(angle = 75, hjust=1, face = "italic", size = 6)) +
  theme(axis.text.y = element_text(size = 6))
  #scale_colour_brewer(palette=color_palette)
total_change
```

### Plot 3

**I also want to see a distinction among provinces** 

_Pls note that: The data only covers May to Nov_
```{r prov-change, include=TRUE}

prov_change <- ggplot(df_long, aes(x=monyear, y=rain_fall)) + 
  #geom_line() +
  geom_point(size=1.5, shape=20) +
  #geom_vline(xintercept=which(as.month(df_long$monyear), nchar(df_long$monyear)-1, nchar(df_long$monyear))=="05"), color = "red", linetype="dashed", size=1.5) +
  facet_wrap(~Province, ncol=3) +
  ggtitle("Precipitation in different provinces over the months and years - A view on variability") +
  theme_grey() +
  labs(y = "Precipitation", colour="Province") +
  theme(axis.title.x = element_blank()) +
  theme(axis.title.y = element_text(size = 8)) +
  theme(axis.text.x = element_text(angle = 75, hjust=1, face = "italic", size = 6)) +
  theme(axis.text.y = element_text(size = 6))

prov_change
```


## Plot 4 - Combination of plot 1&2
```{r}
# color_palette <- c("#000000","#004949","#009292","#ff6db6","#ffb6db",
#  "#490092","#006ddb","#b66dff","#6db6ff","#b6dbff",
#  "#920000","#924900","#db6d00","#24ff24","#ffff6d")

combine <- ggplot(df_long, aes(x=Year, y=rain_fall, colour=Province)) + 
  geom_line() +
  geom_line(data=df_long_cal) +
  #geom_point(size=1.5, shape=20) +
  #geom_hline(yintercept=100, color = "red", size=1.5) +
  facet_wrap(~month, ncol=7) +
  ggtitle("How does rainfall change over the years? - A view on climate change") +
  #theme(text = element_text(size = 16, family="Helvetica", face="italic")) + 
  theme_grey() +
  labs(y = "Precipitation", colour="Province") +
  theme(axis.title.x = element_text(size = 8)) +
  theme(axis.title.y = element_text(size = 8)) +
  theme(axis.text.x = element_text(size = 6)) +
  theme(axis.text.y = element_text(size = 6))
  #scale_colour_brewer(palette=color_palette)
combine
```

## Next step
**1. Map with the reality: Is there any natural events to back up my findings? Eg. Typhoon, Flood, Drought, etc.**

**2. I am curious to know the correlation between rainfall and the gap in highest and lowest quintile
However, I face some challenges:**
*- Quintile income index is annual, while rainfall data is monthly. I will need to choose the representative figure for rainfall index. Should it be mean, median...?*

*- Rainfall data only covers from May to Nov. Would this be sufficient for me infer?* 

**Therefore, I decided to leave this as a question for further investigation.**

```{r income-gap, eval=FALSE, include=FALSE}
prov_long$index_type <- substr(prov_long$index, 1, 6)
income_gap <- prov_long[prov_long$index_type=="income", ]

income_gap$Year_int <- as.integer(income_gap$year)
join <- inner_join(income_gap, df_long, by = c("Year_int" = "Year",
                                               "province" = "Province")
)


```

